### 2018-03-20 writeup/paper.tex
Updated paper

%%% LaTeX Template: Two column article
%%%
%%% Source: http://www.howtotex.com/
%%% Feel free to distribute this template, but please keep to referal to http://www.howtotex.com/ here.
%%% Date: February 2011

%%% Preamble
\documentclass[	DIV=calc,%
			paper=a4,%
			fontsize=11pt,%
			twocolumn]{scrartcl}					% KOMA-article class

\usepackage{lipsum}								% Package to create dummy text

\usepackage[english]{babel}							% English language/hyphenation
\usepackage[protrusion=true,expansion=true]{microtype}			% Better typography
\usepackage{amsmath,amsfonts,amsthm}					% Math packages
\usepackage[pdftex]{graphicx}							% Enable pdflatex
\usepackage[svgnames]{xcolor}							% Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}	% Custom captions under/above floats
\usepackage{epstopdf}								% Converts .eps to .pdf
\usepackage{subfig}								% Subfigures
\usepackage{booktabs}								% Nicer tables
\usepackage{fix-cm}								% Custom fontsizes

%%% Custom sectioning (sectsty package)
\usepackage{sectsty}								% Custom sectioning (see below)
\allsectionsfont{%									% Change font of al section commands
	\usefont{OT1}{phv}{b}{n}%						% bch-b-n: CharterBT-Bold font
}

\sectionfont{%									% Change font of \section command
	\usefont{OT1}{phv}{b}{n}%						% bch-b-n: CharterBT-Bold font
}

%%% Headers and footers
\usepackage{fancyhdr}								% Needed to define custom headers/footers
\pagestyle{fancy}								% Enabling the custom headers/footers
\usepackage{lastpage}

% Header (empty)
\lhead{}
\chead{}
\rhead{}
% Footer (you may change this to your own needs)
\lfoot{\footnotesize \texttt{ECS 272} \textbullet ~Visualization}
\cfoot{}
\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}}		% "Page 1 of 2"
\renewcommand{\headrulewidth}{0.0pt}
\renewcommand{\footrulewidth}{0.4pt}

%%% Creating an initial of the very first character of the content
\usepackage{lettrine}
\newcommand{\initial}[1]{%
     \lettrine[lines=3,lhang=0.3,nindent=0em]{
     				\color{Black}
     				{\textsf{#1}}}{}}

%%% Title, author and date metadata
\usepackage{titling}								% For custom titles

\newcommand{\HorRule}{\color{Black}%					% Creating a horizontal rule
\rule{\linewidth}{1pt}%
}
%%begin novalidate
\pretitle{\vspace{-30pt} \begin{flushleft} \HorRule 
\fontsize{50}{50} \usefont{OT1}{phv}{b}{n} \color{Black} \selectfont 
}
\title{Outlier-preserving Focus and Context Visualization for Parallel Coordinates}		% Title of your article goes here
\posttitle{\par\end{flushleft}\vskip 0.5em}

\preauthor{\begin{flushleft}
\large \lineskip 0.5em \usefont{OT1}{phv}{b}{sl} \color{Black}}
\author{Vincent Yang, Bradley Wang }			% Author name goes here
\postauthor{\footnotesize \usefont{OT1}{phv}{m}{sl} \color{Black} 
					University of California, Davis 		% Institution of author
\date{}										% No date
\par\end{flushleft}\HorRule}
%%end novalidate

%%% Begin document
\begin{document}
\maketitle
\thispagestyle{fancy} 								% Enabling the custom headers/footers for the first page

\section {Goal and Approach}
Our goal in this project was to create a parallel coordinates system that would make outliers and primary trends extremely obvious for the viewer. The system should scale well with large datasets and be an effective tool for finding interesting relationships or outliers. 
Our approach is to implement the same ideas as from the paper. To scale with large datasets, we would simplify large datasets down to a normalized set of trends and outliers, then display these trends/outliers instead. Clustering would enable the user to find unique/interesting trends via. interaction, although this is a supplementary feature.

\section {Our Experience}

\subsection {Positives}
This paper was interesting because it has a new approach to parallel coordinates that we haven't seen before. This was also our first time implementing a visualization paper, 
which gave us insight into how difficult it is to create these systems. Far more effort went into figuring out how to use which tools, which I didn't expect. Next, the paper's approach
and pictures also gave us a strong idea on the benefits of their method. I learned alot on how data may be processed before displaying to the user. We also learned about
how people in visualization have to think about the reality of their visualization methods -- that computing might take a long time or the screen might not be large enough.

\subsection {Negatives}
We found the paper a little bit difficult to make sense of. The title of the paper implied that it would focus on outlier and focus, but it didn’t distinguish the difference between outlier and focus particularly well. This made working on the implementation rather difficult, as we had to read significantly more background literature than previously expected. 
We also found that the paper was somewhat misleading. It promotes efficiency via calculating bucket frequency to accentuate features with parallel coordinates plots. A large part of the paper focuses on the speed benefits of this approach. They also mention that buckets should be calculated between pairs of axes, implying that you would only need to calculate $n-1$ axes pairs where $n$ is the number of axes. However, this is not true at all. Later on in the paper, they bring up the point that if you actually want to be able to reorder the axes (a common feature in parallel coordinates), then you would have to calculate frequencies for every possible combination of axes. This is far more than before! With this, we lose much of the speedup we previously obtained with their method. 
Initially, we implemented the preprocessing by using JavaScript, but found that using Python to precompute the buckets as opposed to computing on page-load made far more sense for the user. With this, the user could simply load data as opposed to processing then loading. 
Some challenges I faced with precomputing was figuring out how to use gaussian smoothing. In the paper, they simply mentioned that it was a common technique in visualization then linked to an early paper regarding gaussian smoothing. This helped very little with my understanding for what to do. I ended up finding blog posts teaching me how to do 2D gaussian smoothing. In the current implementation, we had to look at various smoothnesses dataset by dataset to determine which alpha value (how much to smooth) for each dataset. From what we could find, the proper value should be the standard deviation. Yet with many datasets, we would get completely uniform data, making the data too smooth. In their implementation, they likely found a better way to run smoothing for displaying/calculating the focus. 
Another issue with focus is that they never described how the user would interact with their system to specify which areas to focus on. In traditional parallel coordinates, the user may “scrub” an axis to indicate that they want to look at a subset of data where the values are within a certain range on a given axis. Yet in this implementation, this functionality doesn’t seem to work -- rendering the original values for the user to see would somewhat defeat the purpose of displaying trends and outliers for extremely large datasets. Scrubbing for finding trends or outliers doesn’t make sense either, since the user should already be able to see this clearly.

\section {Paper Information}
TBD

\section {System Differences}
Our implementation allows for the user to specify how many bins they want. The paper doesn’t cover details of how the user interacts with their system, so we decided to simply add a command line argument to the precomputation script. This is accessible by running python3 bin.py n, where n is the number of bins desired per axis. 
Bradley talk about alpha blending here.

\subsection {Datasets Chosen}
The paper’s examples used a CFD simulation dataset (mixture of two fluids) and remote sense data. We were unable to find the original datasets, which made it difficult for us to compare our implementations. 
We chose the international census data to emphasize the speed improvements with their method. The census data is nearly a gigabyte and has over 100 axes, which makes it interesting for us to observe. This dataset was released by the US Census Bureau in 1950 and has its projections through to 2050. Specifically, the original data used to generate this dataset came from observing cities with over 5000 people. We looked specifically at midyear population age by country. 
Our other datasets are smaller. As the number of axes increased, we had a drastic increase in time spent computing. However, with adding more rows, we didn’t observe a similar slowdown. This is likely because adding values to buckets runs in linear time but adding an axis requires that we calculate more buckets for every existing axis. \\

For all of our datasets, we made sure to either choose datasets without numerical categorical variables or to remove those columns. Some useful work that could be done here would be to
map categorical variables to regular intervals, but this would not work well with their approach of creating trends, as the order of the trends would have no inherent importance
but this would affect our trend detection. Because of this, it seems the specific usecase for their approach is actually quite small.

\section {Case Study/Analysis Results}
TBD (timing test results)

\section {Strengths and Limitations}
A key strength of this method is that it scales well for large data sets. Since we simply represent the count in buckets then normalize these buckets, the original size of the dataset has very little influence on that of the precomputed data. 
A limitation of this method and parallel coordinates overall is that it is primarily designed for numerical variables. As such, we can’t work with many datasets. We also can’t represent time series or continuous data well, since we bind them to their closest bucket ranges. 
This work could be furthered by diving deeper into how buckets should be created. Bucket creation is core to their paper and may drastically affect the visualization. We defaulted to using 10 buckets per axis, but this may result in a loss of detail for the final visualization. 
Another idea for future work would be to improve the outlier detection. The outlier detection works by scanning buckets that are at least 10\% full. Changing this barrier would change the outliers visible for the user.

% Including the bibliography
\bibliographystyle{plain}
%\nocite{*}
\bibliography{paper}

\end{document}

### 2018-03-20 writeup/paper.tex
Updated paper

@@
 The system should scale well with large datasets and be an effective tool for finding interesting relationships or outliers.
-Our approach is to implement the same ideas as from the paper.
-To scale with large datasets, we would simplify large datasets down to a normalized set of trends and outliers, then display these trends/outliers instead.
-Clustering would enable the user to find unique/interesting trends via.
-interaction, although this is a supplementary feature.
+This would allow us
+to scale with large datasets.
+
+As for personal goals, this is our first time implementing a visualization paper, so we
+hope to learn about what goes into a research paper furthering visualization. 
+Thus far in class, we have only seen the end results and learned about using others' tools,
+but haven't created our own.
+
+Our approach is to for precompute the data we need. This would involve (a) binning the data, (b) figuring out 
+where the main trends are, (c) figuring out where the outliers are, and (d) figuring out where the clusters are. 
+For binning the data, we will look at every combination of 2 axes just as they specify in the paper. Of each pair,
+we will create a table of bins representing the frequency of certain bin combinations. With these, we will take the
+more populated bins and call them our trends, then finding the outliers by using median and isolation filters.
+
+Clustering will involve using gaussian smoothing then iteratively moving from more populated bins
+to less populated bins to determine where interesting clusters might lie.
+
+Once we have precomputed this data, we will plot it on our system with d3 and webgl. d3 creates
+the axes and labels, and supports the ability to move axes around. Webgl will display our polygons for
+representing trends as well as drawing the outliers. This would also take care of the textures for us.
+
+To run the precomputing script on data.csv with 10 bins per axis, run \texttt{python3 bin.py data 10}. The input
+data should have no quotes and end in \texttt{.csv}.
 
@@
 \subsection {Negatives}
-We found the paper a little bit difficult to make sense of.
+We found the paper quite difficult to make sense of.
 The title of the paper implied that it would focus on outlier and focus, but it didn’t distinguish the difference between outlier and focus particularly well.
@@
 
-\section {Paper Information}
-TBD
-
 \section {System Differences}
@@
 Bradley talk about alpha blending here.
+
+Our implementation also has axis labels so the user can see exactly where trends/outliers lie. In their implementation,
+they have labels at the very top and bottom, but nowhere in between.
+
+They discuss using the GPU to optimize rendering, but we don't interact with the GPU. We wanted to focus on 
+getting a working implementation and observing the benefits of binning.
 
@@
 
+The list of datasets we chose are BudgetFood, EuStockMarkets, and MidyearPopulationCensus. The largest dataset is MidyearPopulationCensus, which is over 800 mb. We
+cleaned out all of the datasets by converting male/female values to 0 and 1, and removing non-numerical columns.
+
 \section {Case Study/Analysis Results}
@@
 \section {Strengths and Limitations}
-A key strength of this method is that it scales well for large data sets.
-Since we simply represent the count in buckets then normalize these buckets, the original size of the dataset has very little influence on that of the precomputed data.
+A key strength of this method is that there is far less data to plot at a time.
+When we tried to render traditional parallel coordaintes with some gigabyte datasets, it took almost a minute just to load all of the data onto
+the browser, then more to draw. Although it took a while to precompute our implementation, it rendered far faster. Additionally, it would be fully possible to parallelize our
+precomputation to enable the user to spend even less time in this step.
+
 A limitation of this method and parallel coordinates overall is that it is primarily designed for numerical variables.

### 2018-03-20 simple-parcoord/bin.py
Created simple parallel coordinates system

@@ -1,0 +1,513 @@
+import sys
+import json
+import csv
+import matplotlib.pyplot as pl
+import numpy as np
+import scipy as sp
+import scipy.ndimage
+from statistics import median
+
+def no_str_keys(d):
+    if not d:
+        return d
+    for k in d.keys():
+        if type(k) is not str:
+            try:
+                d[str(k)] = d[k]
+            except:
+                try:
+                    d[repr(k)] = d[k]
+                except:
+                    pass
+            del i[k]
+    return(d)
+
+def writejson(filename, bins, outliers, clusters, num_dimensions):
+    l = [[None for i in range(num_dimensions)] for j in range(num_dimensions)]
+
+    #with open(filename, 'w') as out:
+    #    for i in range(num_dimensions):
+    #        for j in range(num_dimensions):
+    #            di = {}
+
+    #            di['bins'] = (bins[i][j]) # num_bins x num_bins; each has stuff
+    #            di['outliers'] = ((outliers[i][j]))
+    #            di['clusters'] = (clusters)
+
+    #            l[i][j] = di
+    #    s = json.dumps(l, separators=(',', ':'))
+    #    out.write(s)
+    for i in range(num_dimensions):
+        for j in range(num_dimensions):
+            di = {}
+
+            di['bins'] = (bins[i][j]) # num_bins x num_bins; each has stuff
+            di['outliers'] = ((outliers[i][j]))
+            di['clusters'] = (clusters)
+
+            l[i][j] = di
+    return(l)
+
+
+"""
+Write out 2d axis data 
+"""
+def output(filename, bins):
+    with open(filename, 'w') as out:
+        for i in range(len(bins)):
+            out.write(str(bins[i]))
+            out.write('\n')
+
+def getNeighbors(x,y):
+    if x == 0 and y == 0:
+        return [(0,1), (1,1), (1,0)]
+    elif x == 0:
+        return [(0,y+1), (0,y-1), (1,y+1), (1,y), (1,y-1)]
+    elif y == 0:
+        return [(x,1), (x+1,0), (x-1,0), (x+1,1), (x-1,1)]
+    else:
+        return([(x,y+1), (x,y-1), (x+1,y+1), (x+1,y-1), (x-1,y+1), (x-1,y-1), (x+1, y), (x-1, y)])
+
+"""
+Cluster data points for user interaction with focus
+"""
+def cluster(bins, num_bins):
+    # smooth out the data with gaussian filter
[…]

### 2018-03-20 simple-parcoord/parcoord-script.js

@@ -1,0 +1,278 @@
+var margin = {top: 66, right: 110, bottom: 20, left: 188},
+    width = document.body.clientWidth - margin.left - margin.right,
+    height = 340 - margin.top - margin.bottom,
+    innerHeight = height - 2;
+
+var devicePixelRatio = window.devicePixelRatio || 1;
+
+var color = d3.scaleOrdinal()
+    .domain(["Radial Velocity", "Imaging", "Eclipse Timing Variations", "Astrometry", "Microlensing", "Orbital Brightness Modulation", "Pulsar Timing", "Pulsation Timing Variations", "Transit", "Transit Timing Variations"])
+    .range(["#DB7F85", "#50AB84", "#4C6C86", "#C47DCB", "#B59248", "#DD6CA7", "#E15E5A", "#5DA5B3", "#725D82", "#54AF52", "#954D56", "#8C92E8", "#D8597D", "#AB9C27", "#D67D4B", "#D58323", "#BA89AD", "#357468", "#8F86C2", "#7D9E33", "#517C3F", "#9D5130", "#5E9ACF", "#776327", "#944F7E"]);
+
+var types = {
+    "Number": {
+        key: "Number",
+        coerce: function(d) { return +d; },
+        extent: d3.extent,
+        within: function(d, extent, dim) { return extent[0] <= dim.scale(d) && dim.scale(d) <= extent[1]; },
+        defaultScale: d3.scaleLinear().range([innerHeight, 0])
+    },
+    "String": {
+        key: "String",
+        coerce: String,
+        extent: function (data) { return data.sort(); },
+        within: function(d, extent, dim) { return extent[0] <= dim.scale(d) && dim.scale(d) <= extent[1]; },
+        defaultScale: d3.scalePoint().range([0, innerHeight])
+    },
+    "Date": {
+        key: "Date",
+        coerce: function(d) { return new Date(d); },
+        extent: d3.extent,
+        within: function(d, extent, dim) { return extent[0] <= dim.scale(d) && dim.scale(d) <= extent[1]; },
+        defaultScale: d3.scaleTime().range([innerHeight, 0])
+    }
+};
+
+var printBin = function(bins) {
+    for (let i = 0; i < bins.length; i++) {
+        for (let j = 0; j < bins[i].length; j++){
+            console.log(bins[i][j])
+        }
+        console.log('');
+    }
+
+}
+
+var dimensions = [ ];
+
+
[…]

### 2018-03-20 src/projFinal.js
Added WebGL rendered back.

@@ -43,6 +43,11 @@ function drawParallelCoordinates() {
         .attr("height", height);
 
     pcSize = [width - chartMargins.left * 2, height - chartMargins.top * 2];
+    let xrange = [];
+    for (let i = 0; i < pcDimensions.length; ++i) {
+        xrange.push(i * pcSize[0] / (pcDimensions.length - 1));
+    }
+    pcXScale.domain(pcDimensions.map(extractId)).range(xrange);
 
     const dpr = window.devicePixelRatio || 1;
 
@@ -56,7 +61,11 @@ function drawParallelCoordinates() {
     wglSetGLObj(canvas.node().getContext("webgl2"));
 
     initDraw();
-    wglDrawFrame();
+
+    wglSetNumPanels(3);
+
+    wglRedrawContext(contextData, pcDimensions.map(extractId));
+    wglDrawFrame(pcDimensions.map(positionNormalized));
 
     drawParallelCoordinatesAxes();
 }
@@ -71,11 +80,6 @@ function drawParallelCoordinatesAxes() {
         .attr("transform", "translate(" + (chartMargins.left) + ", " + (chartMargins.top) + ")");
 
 
-    let xrange = [];
-    for (let i = 0; i < pcDimensions.length; ++i) {
-        xrange.push(i * pcSize[0] / (pcDimensions.length - 1));
-    }
-    pcXScale.domain(pcDimensions.map(extractId)).range(xrange);
 
     for (const dim of pcDimensions) {
         dim.yScale.range([pcSize[1], 0]);
@@ -103,7 +107,12 @@ function drawParallelCoordinatesAxes() {
                 pcDimensions.sort(function(a, b) {return position(a) - position(b);});
                 pcXScale.domain(pcDimensions.map(extractId));
                 axes.attr("transform", function(d) {return "translate(" + position(d) + ")"; });
+                if (Math.abs(d3.event.x - this.__origin__) > pcSize[0]/(pcDimensions.length - 1)) {
+                    this.__origin__ = d3.event.x;
+                }
                 // brush_count++;
+                wglRedrawContext(contextData, pcDimensions.map(extractId));
+                wglDrawFrame(pcDimensions.map(positionNormalized));
                 this.__dragged__ = true;
                 // console.log("drag");
             })
@@ -111,7 +120,10 @@ function drawParallelCoordinatesAxes() {
                 if (!this.__dragged__) {
                     // No movement, basically clicked, do something?
                 } else {
[…]

### 2018-03-20 src/textrans.vsf.js

@@ -1,0 +1,13 @@
+const textransVS = `
+precision mediump float;
+
+attribute vec2 pos;
+attribute vec2 tC;
+
+varying vec2 posOut;
+
+void main() {
+    posOut = tC;
+    gl_Position = vec4(pos, 0.0, 1.0);
+}
+`;

### 2018-03-20 src/texture.fsf.js

@@ -1,0 +1,11 @@
+const textureFS = `
+precision mediump float;
+
+varying vec2 posOut;
+
+uniform sampler2D samp;
+
+void main() {
+    gl_FragColor = texture2D(samp, posOut);
+}
+`;

### 2018-03-20 src/wglparcoord.js

@@ -1,6 +1,12 @@
+const OUTLIER_INTENSITY = 0.35;
+
 let gl;
 
-let shaderProg;
+let colorShaderProg;
+let textureShaderProg;
+
+let texCoordsVBO;
+let panelCoordsVBO;
 
 let parcoordsPanels = [];
 
@@ -19,76 +25,223 @@ function wglSetNumPanels(numPanels) {
             for (const panel of parcoordsPanels) {
                 gl.deleteTexture(panel.texture);
                 gl.deleteFramebuffer(panel.framebuffer);
-                gl.deleteBuffer(panel.vbo);
+                gl.deleteBuffer(panel.trendsVBO);
+                gl.deleteBuffer(panel.outliersVBO);
             }
         }
         initPanels(numPanels, gl.canvas.clientHeight);
+    } else {
+        for (const panel of parcoordsPanels) {
+            panel.invalid = true;
+        }
     }
+}
+
+function wglInvalidatePanel(index) {
+    parcoordsPanels[index].invalid = true;
 }
 
 function initDraw() {
     if (!gl) return;
 
-
     resize();
-    // TODO load shaders
-    shaderProg = {program:linkShaderProgram(transVS, colorFS)};
-    // const vertShad = gl.loadShader(gl.VERTEX_SHADER, vsSource);
 
-    gl.clearColor(0.0,0.0,0.0,0.5); // Set canvas to transparent and blank when clearing.
+    gl.clearColor(0.0,0.0,0.0,0.6); // Set canvas to transparent and blank when clearing.
     gl.disable(gl.DEPTH_TEST);
+    gl.enable(gl.BLEND);
 
-    gl.blendFuncSeparate(gl.SRC_ALPHA, gl.ONE_MINUS_SRC_ALPHA, gl.ONE, gl.ONE_MINUS_SRC_ALPHA);
+    gl.blendFunc(gl.ONE, gl.ONE_MINUS_SRC_ALPHA);
+
+    let shadProg = linkShaderProgram(transVS, colorFS);
+    colorShaderProg = {
+        program: shadProg,
+        attribLocations: {
+            pos: gl.getAttribLocation(shadProg, "pos"),
+            alpha: gl.getAttribLocation(shadProg, "alpha")
+        }
+    };
+    shadProg = linkShaderProgram(textransVS, textureFS);
+    textureShaderProg = {
+        program: shadProg,
+        attribLocations: {
+            pos: gl.getAttribLocation(shadProg, "pos"),
[…]

### 2018-03-20 writeup/paper.tex

@@
 To run the precomputing script on data.csv with 10 bins per axis, run \texttt{python3 bin.py data 10}. The input
-data should have no quotes and end in \texttt{.csv}.
+data should have no quotes and end in \texttt{.csv}. This will also create array.png and array2.png. These show examples of the 
+data before and after gaussian filtering.
+
+\begin{figure}
+  \caption{Our Implemented System}
+  \centering
+    \includegraphics[width=9cm]{alc-opf.png}
+\end{figure}
+
+\begin{figure}
+  \caption{Traditional Parallel Coordinates}
+  \centering
+    \includegraphics[width=9cm]{age-alc-orig.png}
+\end{figure}
+
+From these pictures, we can see the difference between our implemented system and a traditional
+parallel coordinates plot. Even on a small dataset with only several hundred rows, 
+it is easier to notice on the lower right that there are many people in lower ends of both Dalc and Walc.
+
+We can also find the outliers easier, which are clearly in the upper right for our implemented system. 
+It is difficult to see the difference between outliers and trends in conventional parallel coordinates.
 
@@
 We found the paper quite difficult to make sense of.
-The title of the paper implied that it would focus on outlier and focus, but it didn’t distinguish the difference between outlier and focus particularly well.
+The title of the paper implied that it would focus on outlier and focus, but it didn't distinguish the difference between outlier and focus particularly well.
 This made working on the implementation rather difficult, as we had to read significantly more background literature than previously expected.
@@
 In traditional parallel coordinates, the user may “scrub” an axis to indicate that they want to look at a subset of data where the values are within a certain range on a given axis.
-Yet in this implementation, this functionality doesn’t seem to work -- rendering the original values for the user to see would somewhat defeat the purpose of displaying trends and outliers for extremely large datasets.
-Scrubbing for finding trends or outliers doesn’t make sense either, since the user should already be able to see this clearly.
+Yet in this implementation, this functionality doesn't seem to work -- rendering the original values for the user to see would somewhat defeat the purpose of displaying trends and outliers for extremely large datasets.
+Scrubbing for finding trends or outliers doesn't make sense either, since the user should already be able to see this clearly.
 
@@
 Our implementation allows for the user to specify how many bins they want.
-The paper doesn’t cover details of how the user interacts with their system, so we decided to simply add a command line argument to the precomputation script.
+The paper doesn't cover details of how the user interacts with their system, so we decided to simply add a command line argument to the precomputation script.
 This is accessible by running python3 bin.py n, where n is the number of bins desired per axis.
@@
 \subsection {Datasets Chosen}
-The paper’s examples used a CFD simulation dataset (mixture of two fluids) and remote sense data.
+The paper's examples used a CFD simulation dataset (mixture of two fluids) and remote sense data.
 We were unable to find the original datasets, which made it difficult for us to compare our implementations.
@@
 As the number of axes increased, we had a drastic increase in time spent computing.
-However, with adding more rows, we didn’t observe a similar slowdown.
+However, with adding more rows, we didn't observe a similar slowdown.
 This is likely because adding values to buckets runs in linear time but adding an axis requires that we calculate more buckets for every existing axis.
@@
 \section {Case Study/Analysis Results}
-TBD (timing test results)
+\begin{tabular}{ |p{2cm}||p{1cm}||p{1cm}||p{1.2cm}|  }
+ \hline
+ \multicolumn{4}{|c|}{Time in Sec.} \\
+ \hline
+ Data& Orig. Time& Our Time & Rows\\
+ \hline
+ creditdefault& 13 & 8 & 10001\\
+ quakes&   3 & 10 & 1001\\
+ census &- &32min & 3058281\\
+ EuStock &3.1 &8 & 1861\\
+ age-alc &<1 &<1 & 396\\
+ \hline
+\end{tabular}
+
+The strength of this system is clearly shown in the census data. With the normal 
+parallel coordinates implementation, the d3 couldn't even handle holding the data in 
+memory and threw an allocation size overflow error. Admittedly, this dataset also
+took over half an hour to precompute. The long precomputing time is likely due to
+the high number of axes -- there are over 100 axes in this dataset, meaning 
+our precomputation calculated $100^2$ possible combinations of axes.
+
+With most of the small datasets of only several thousand rows, rendering the data
+was extremely quick with traditional parallel coordinates. This is as expected, as 
+there is minimal improvement, if any, from counting frequencies when there isn't much to draw anyway.
 
@@
 A key strength of this method is that there is far less data to plot at a time.
-When we tried to render traditional parallel coordaintes with some gigabyte datasets, it took almost a minute just to load all of the data onto
+When we tried to render traditional parallel coordinates with some gigabyte datasets, it took almost a minute just to load all of the data onto
 the browser, then more to draw. Although it took a while to precompute our implementation, it rendered far faster. Additionally, it would be fully possible to parallelize our
@@
 
+As we can see in this image, it is rather difficult to make sense of what is in the original data, as
+there are so many data points on the right. This is the credit default dataset, which shows
+the attributes of people who default or don't default on their credit card payments. Just rendering
+this dataset which has 10,000 lines takes around 12 seconds. 
+\includegraphics[width=9cm]{credit-default-parcoord.png}
+
 A limitation of this method and parallel coordinates overall is that it is primarily designed for numerical variables.
-As such, we can’t work with many datasets.
-We also can’t represent time series or continuous data well, since we bind them to their closest bucket ranges.
+As such, we can't work with many datasets.
+We also can't represent time series or continuous data well, since we bind them to their closest bucket ranges.
 This work could be furthered by diving deeper into how buckets should be created.
@@
 \bibliography{paper}
-
 \end{document}

### 2018-03-20 src/projFinal.js
Dynamic panel init

@@ -62,7 +62,7 @@ function drawParallelCoordinates() {
 
     initDraw();
 
-    wglSetNumPanels(3);
+    wglSetNumPanels(pcDimensions.length - 1);
 
     wglRedrawContext(contextData, pcDimensions.map(extractId));
     wglDrawFrame(pcDimensions.map(positionNormalized));

### 2018-03-20 simple-parcoord/bin.py
Cleanup

@@ -422,14 +422,12 @@ if __name__ == '__main__':
         num_bins = int(sys.argv[2])
     else:
         num_bins = 10
-    print('num_bins is ' + str(num_bins))
     dimensions = text[0]
     num_dimensions = len(dimensions)
     axesbins = [[None for i in range(num_dimensions)] for i in range(num_dimensions)]
     clusterbins = [[None for i in range(num_dimensions)] for i in range(num_dimensions)]
     # Given two axes i and j, axesoutliers[i][j] holds a list of (val1, val2) that are outliers. 
     axesoutliers = [[None for i in range(num_dimensions)] for i in range(num_dimensions)]
-    #print(dimensions)
     
     # find max and min for each dimension; max[i] holds the max for the i'th dimension, or dimensions[i]
     max_dim, min_dim = getMaxMin(text)
@@ -454,7 +452,6 @@ if __name__ == '__main__':
                 min_freq = min(min_freq, bins[bucket1][bucket2])
                 total_max_freq = max(max_freq, total_max_freq)
 
-            #print('Comparing ' + str(dimensions[i]) + ' and ' + str(dimensions[j]))
             # Create outliers
             isolation_bins = isolationFilter(int(max_freq * 0.1), bins)
             median_bins = medianFilter(int(max_freq * 0.1), bins)
@@ -482,7 +479,6 @@ if __name__ == '__main__':
 
             # clustering
             c = cluster(bins, num_bins)
-            print(c)
             # invert clusters for formatting. Make it key(cluster#) -> array of locations
             cluster_d = {}
             for k in c.keys():
@@ -491,7 +487,6 @@ if __name__ == '__main__':
                 else:
                     cluster_d[c[k]] = [k]
                 
-            print(cluster_d)
             clusterbins[i][j] = cluster_d
 
             axesoutliers[i][j] = outliers

### 2018-03-20 simple-parcoord/parcoord-script.js

@@ -47,7 +47,12 @@ var dimensions = [ ];
 
 
 //d3.csv("planets.csv", function(error, data) {
-d3.csv("credit-default.csv", function(error, data) {
+//d3.csv("quakes.csv", function(error, data) {
+
+//d3.csv("cleaned-midyear-population-age-country-code.csv", function(error, data) {
+d3.csv("age-alc.csv", function(error, data) {
+//d3.csv("EuStockMarkets.csv", function(error, data) {
+//d3.csv("credit-default.csv", function(error, data) {
 //d3.csv("cleaned-midyear-population-age-country-code.csv", function(error, data) {
     if (error) throw error;
 

### 2018-03-20 simple-parcoord/runner.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+date -u 
+python3.6 bin.py credit-default 10
+date -u

### 2018-03-20 src/bin.py

@@ -406,29 +406,28 @@ def medianFilter(threshold, grid):
 
     return(res)
 
-filename = 'age-alc'
-#filename = 'test'
-with open (filename + '.csv', 'r') as f:
-    reader = csv.reader(f, delimiter = ',', quoting=csv.QUOTE_NONNUMERIC)
-    text = []
-    for row in reader: 
-        text.append(row)
-
 if __name__ == '__main__':
-    # initialize relevant variables
-    print(sys.argv)
     if len(sys.argv) > 1:
-        num_bins = int(sys.argv[1])
+        filename = sys.argv[1]
+    else:
+        filename = 'age-alc'
+    with open (filename + '.csv', 'r') as f:
+        reader = csv.reader(f, delimiter = ',', quoting=csv.QUOTE_NONNUMERIC)
+        text = []
+        for row in reader: 
+            text.append(row)
+    
+    # initialize relevant variables
+    if len(sys.argv) > 2:
+        num_bins = int(sys.argv[2])
     else:
         num_bins = 10
-    print('num_bins is ' + str(num_bins))
     dimensions = text[0]
     num_dimensions = len(dimensions)
     axesbins = [[None for i in range(num_dimensions)] for i in range(num_dimensions)]
     clusterbins = [[None for i in range(num_dimensions)] for i in range(num_dimensions)]
     # Given two axes i and j, axesoutliers[i][j] holds a list of (val1, val2) that are outliers. 
     axesoutliers = [[None for i in range(num_dimensions)] for i in range(num_dimensions)]
-    #print(dimensions)
     
     # find max and min for each dimension; max[i] holds the max for the i'th dimension, or dimensions[i]
     max_dim, min_dim = getMaxMin(text)
@@ -453,7 +452,6 @@ if __name__ == '__main__':
                 min_freq = min(min_freq, bins[bucket1][bucket2])
                 total_max_freq = max(max_freq, total_max_freq)
 
-            #print('Comparing ' + str(dimensions[i]) + ' and ' + str(dimensions[j]))
             # Create outliers
             isolation_bins = isolationFilter(int(max_freq * 0.1), bins)
             median_bins = medianFilter(int(max_freq * 0.1), bins)
@@ -481,7 +479,6 @@ if __name__ == '__main__':
 
             # clustering
             c = cluster(bins, num_bins)
-            print(c)
             # invert clusters for formatting. Make it key(cluster#) -> array of locations
             cluster_d = {}
[…]

### 2018-03-20 src/parcoord-script.js

@@ -158,81 +158,6 @@ d3.csv("test.csv", function(error, data) {
         dim.scale.domain(dim.domain);
     });
 
-    // Do data processing here
-    console.log(data);
-    console.log(Object.keys(data[0]));
-    const numBins = 10;
-    const dims = Object.keys(data[0]);
-    const numDimensions = dims.length;
-
-
-    // create 2d array to represent every possible combination.
-    const comb = new Array(numDimensions);
-    for (let i = 0; i < numDimensions; i++) {
-        comb[i] = new Array(numDimensions);
-        console.log(Object.keys(data[0])[i]);
-    }
-
-    // find min, max for each
-    const minvals = new Array(numDimensions);
-    const maxvals = new Array(numDimensions);
-    for (let i = 0; i < numDimensions; i++) {
-        minvals[i] = data[0][dims[i]];
-        maxvals[i] = data[0][dims[i]];
-    }
-    for (let i = 0; i < numDimensions; i++) {
-        for (let j = 0; j < data.length; j++) {
-            minvals[i] = Math.min(minvals[i], data[j][dims[i]]);
-            maxvals[i] = Math.max(maxvals[i], data[j][dims[i]]);
-        }
-    }
-    console.log(minvals);
-    console.log(maxvals);
-
-
-    // for each possible combination, create an array of tuples.
-    for (let i = 1; i < numDimensions; i++) {
-        for (let j = i+1; j < numDimensions; j++) {
-            // between each pair of axes, we need to create a 2 by 2 array of numBins
-            const bins = new Array(numBins);
-            for (let k = 0; k < numBins; k++) {
-                bins[k] = Array.apply(null, Array(numBins)).map(Number.prototype.valueOf,0);
-            }
-
-            for (let k = 0; k < data.length; k++) {
-                let val1 = data[k][dims[i]];
-                let val2 = data[k][dims[j]];
-                let bucket1 = Math.trunc((val1 - minvals[i])/(maxvals[i] - minvals[i]) * 10);
-                let bucket2 = Math.trunc((val2 - minvals[j])/(maxvals[j] - minvals[j]) * 10);
-                //console.log('val1: ' + val1 + ', min: ' + minvals[i] + ', max: ' + maxvals[i] + ', bucket: ' + bucket1);
-                //console.log('val2: ' + val2 + ', min: ' + minvals[j] + ', max: ' + maxvals[j] + ', bucket: ' + bucket2);
-                if (bucket1 === numBins) {
[…]

### 2018-03-20 src/time-test/bin.py

@@ -1,467 +1,0 @@
-import json
-import csv
-import matplotlib.pyplot as pl
-import numpy as np
-import scipy as sp
-import scipy.ndimage
-from statistics import median
-
-def writejson(filename, bins, outliers, clusters):
-    l = []
-
-    with open(filename, 'w') as out:
-        for i in range(len(bins)):
-            for j in range(len(bins)):
-                dict = {}
-                dict['bins'] = str(bins[i][j])
-                dict['outliers'] = str(outliers[i][j])
-                dict['clusters'] = str(clusters)
-                #print(dict)
-                #s = json.dumps(dict, separators=(',', ':'))
-                #out.write(s)
-                l.append(dict)
-        s = json.dumps(l, separators=(',', ':'))
-        out.write(s)
-
-"""
-Write out 2d axis data 
-"""
-def output(filename, bins):
-    with open(filename, 'w') as out:
-        for i in range(len(bins)):
-            out.write(str(bins[i]))
-            out.write('\n')
-
-def getNeighbors(x,y):
-    if x == 0 and y == 0:
-        return [(0,1), (1,1), (1,0)]
-    elif x == 0:
-        return [(0,y+1), (0,y-1), (1,y+1), (1,y), (1,y-1)]
-    elif y == 0:
-        return [(x,1), (x+1,0), (x-1,0), (x+1,1), (x-1,1)]
-    else:
-        return([(x,y+1), (x,y-1), (x+1,y+1), (x+1,y-1), (x-1,y+1), (x-1,y-1), (x+1, y), (x-1, y)])
-
-"""
-Cluster data points for user interaction with focus
-"""
-def cluster(bins, num_bins):
-    # smooth out the data with gaussian filter
-    gaus_bins = gaus_filter(bins, num_bins)
-
-    # iteratively go from the most frequent to 10% of the most frequent
-    # when a block may be added to both clusters, add it to the one with the closest 
-    # peak height
-    
-    clusters = {}
-    # find the order of coordinates sorted by frequency
-    l = []
-    for i in range(num_bins):
-        for j in range(num_bins):
-            if gaus_bins[i][j] != 0:
-                l.append((gaus_bins[i][j], (i, j))) # freq, tuple coordinates
-    l.sort(key=lambda x: x[0], reverse=True)
-
-    cluster_map = [[0 for a in range(num_bins)] for b in range(num_bins)]
-    max_freq = l[0][0]
-    for freq, tup in l:
[…]

### 2018-03-20 writeup/paper.tex
Cleanup
clean repo

@@
 
-To run the precomputing script on data.csv with 10 bins per axis, run \texttt{python3 bin.py data 10}. The input
-data should have no quotes and end in \texttt{.csv}. This will also create array.png and array2.png. These show examples of the 
-data before and after gaussian filtering.
+To run the precomputing script on data.csv with 10 bins per axis, run \texttt{python3.6 bin.py data 10}. The input
+data should have no quotes and end in \texttt{.csv}. We require python3.6 for matplotlib, which is needed
+for gaussian filtering. This will also create array.png and array2.png. These show examples of the 
+data before and after gaussian filtering. Alternatively, there are pre-created output.json files by us in /src/precomputed-data. 
+Move those into /src and rename to ``output.json'' to try each out.
 
@@
 This is accessible by running python3 bin.py n, where n is the number of bins desired per axis.
-Bradley talk about alpha blending here.
 
@@
  EuStock &3.1 &8 & 1861\\
- age-alc &<1 &<1 & 396\\
+ age-alc &$<1$ &$<1$ & 396\\
  \hline
